Update and evaluation of the ozone dry deposition in the Oslo~CTM3 v1.0

### 2019-05-23 version_2/response_version_1_rc2.tex

% Authors' response
\documentclass{scrartcl}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
%\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{xcolor}
\usepackage{url}
\usepackage{multirow}
\usepackage{graphicx}
\thispagestyle{empty}

\begin{document}
\section*{Authors' response}
To gmd-2019-21 David Simpson (10 Apr 2019):
We apologize for the late response regarding the comments on our paper. We had to evaluate the major concerns raised in the review and found that we have to revise our model and repeat the model experiments to address all concerns. Nevertheless, we will respond to the questions in more detail in the following.

\section{Major points}
\begin{itemize}
\item {\color{blue} Mosaic formulation:}
  \begin{itemize}
  \item {\color{blue} CTM3 claims to implement a mosaic approach, but instead of calculating deposition
rates over each land-cover, and then aggregating using Eqn.3, CTM3 seems to perform the following steps:
\begin{align}
  \label{eq:ra}
  R_a &= \frac{u_z}{u^2_*}\\
  \label{eq:Gsto}
  \overline{Gs^i_g(z)} &= \Sigma^N_{k=1} f_k \times Gs^i_{g,k}(z)\\
  \label{eq:Gns}
  \overline{Gns^i_g(z)} &= \Sigma^N_{k=1} f_k \times Gns^i_{g,k}(z)
\end{align}
As far as I can tell from the text, $R_a$ is calculated once, with the same value of $u_*$ for all land-covers. The CTM3 $R_b$ calculation seems to also use the same $u_*$ over different land-covers, except over sea where a more sophisticated scheme is used. Equations 5-6 above are equivalent to Eqns (22) and (26) from the manuscript. Now I am puzzled however as to how all this is put together. Do they use:
\begin{equation}
  G_c = \mathrm{LAI} \cdot \overline{Gs^i_g(z)} + \overline{Gns^i_g(z)}
\end{equation}
  }
    The referee is correct regarding the usage of Eq.~(\ref{eq:ra}) and Eq.~(\ref{eq:Gns}). In case of stomatal conductance, we first compute:
    \begin{equation}
       \overline{gs^i_g(z)} = \Sigma^N_{k=1} f_k \times gs^i_{g,k}(z).
    \end{equation}
    This is put together as follows:
    \begin{equation}
      \label{eq:Gsto_corr}
      G_c = \mathrm{LAI} \cdot \overline{gs^i_g(z)} + \overline{Gns^i_g(z)}
    \end{equation}
  \item {\color{blue} In any case, I think this approach has serious problems. Why average first for $G_s$ and then for $G_{ns}$ , when it is the fluxes ($F_k$ , or $V^i_{g,k}(z_\mathrm{ref}) \times \chi^i_\mathrm{avg}(z_\mathrm{ref})$) which need to be averaged? I also do not understand why they would use the same $u_*$ and $R_a$ for all land-covers. I think the authors need to make a case for their approach, or change it.}
    Thank you very much for your detailed account of the \emph{mosaic approach}. We have discussed the concerns raised within this comment and found that we have indeed made a mistake in our implementation of the \emph{mosaic approach} which forces us to revise our model and repeat the model experiments.
  \end{itemize}
\item {\color{blue} Calculation of $R_a$:}
  \begin{itemize}
  \item {\color{blue} $R_a$ in CTM3 appears to be calculated just once, and from a height of 8\,m. This means
that there is no consistency between the $R_a$ term and the underlying surface, which is
clearly wrong. The similarity equation for $R_a$ given in their Eqn~(2) is very standard and has been
used for decades (Garratt 1992). As pointed out by Ref~1, the equation as given has
errors. The correct equation will not give negative values unless presented with wrong
inputs, and I suspect that that is what has happened.}
We have, in fact, not presented "wrong inputs" to the algorithm.
As pointed out by Ref \#1, Eq. (2) in our manuscript is erroneous. The corrected equation (Erisman, 1994) for stable atmospheric conditions:
\begin{equation}
  R_a = \frac{1}{\kappa u_*}\left[{\ln{\left(\frac{z-d}{z_0}\right)}-\Psi_m\left(\frac{z-d}{L}\right)+\Psi_m\left(\frac{z_0}{L}\right)}\right].
  \label{eq:aero_res}
\end{equation}

@@

\item {\color{blue}P8 L18. This is wrong. Nothing in the EMEP model is used to ’mimic the
time lag..’. We use the light function to modify stomatal conductance, as with the
other $f$ factors.}
  We change this inadequate formulation: \emph{"[...] this integrated photon flux is used to modify the stomatal conductance in response to light."}

\item {\color{blue}P9 L20, and Table 1. The consequence of Table 1 is that vegetation at $0.5^\circ\,\mathrm{N}$
will start growing at day 90, whereas those at $0.5^\circ\,\mathrm{S}$ will start on day 272. (By the
way, in EMEP now we use monthly factors from the LPJ-GUESS model to derive
phenology for non-European areas, because of such difficulties with tabulations.)}
  Thank you for the remark. We see this problem. Using an actual land model would solve this, but at that point it is not feasible to integrate such in the Oslo~CTM3.

\item {\color{blue}P9 L18. what do you mean by "surface or 2\,m"? Surface might refer to skin or
  leaf temperature?}

\item {\color{blue}P9 Eqn (26). Say 1st and 2nd, not I. and II.}
  We follow the advise regarding Eq.~(26) and also change Table~1 accordingly.

\item {\color{blue}P9 Eqn (27). This equation is a modification of Erisman’s original (1994) formulation, so explain that.}
  We follow the advise and write: \emph{"The in-canopy resistance $R_\text{inc}$ (Erisman et al., 1994) is then modified with respect to each (vegetated) land type $N$ [...]"} {\bfseries ToDo -- To Amund: Were does this modification come from?}

\item {\color{blue}P10 L12. Be explicit that this statement refers to S2012. The current EMEP
  model uses different heights for e.g. tropical vegetation.}
  We follow this advice and write: \emph{"The vegetation height $h(N, \text{lat})$ as decribed by Simpson et al. (2012) [...]"}

\item {\color{blue}P11 Sect 2.2. I also found this aerosol section confusing. Eqn (30) is from S2012,
and so is the factor 0.008.SAI/10 used in Eqn (33), but here new $a_1$ coefficients
are defined. Did the ’aerosol microphysic model’ referred to also mix equations
in this way? Is there any publication as to the reliability of this method?}
  The $a_1$ coefficients were actually calculated from deposition velocities taken from the Oslo~CTM2 which has been socumented {\bfseries ToDo: Citation of CTM2?}.

\item {\color{blue}P12 Fig.2. I didn’t understand what is being done here. The Figure suggests that
the EMEP scheme has one category for ’Forests, Med. scrub’, whereas S2012
lists 4 types of forest, as well as Mediterranean scrub as a separate ecosystem.
This figure also suggests that EMEP has savanna, which it doesn’t, but do have
many other categories (Table 3 of S2012 lists 16 main categories. The current
EMEP model has 32.)}
  The caption may become clearer after we changed the name of our scheme, so that it can no longer be confused with the actual EMEP model. The purpose of the mapping is to project the more detailed categories for stomatal conductance to the 10 ground-surface resistance categories (Table~S19 in S2012 supplement).

\item {\color{blue}P12 L13. Again, the current EMEP model is not eurocentric, and uses global
  phenology calculations.}
  We follow the advice and change the text accordingly: \emph{"Since the eurocentric parameterization of SGS and EGS in Simpson et al. (2012) is not applicable [...]"}

\item {\color{blue}P13 Sect 2.3.2. The initial lines (14-16) are hard to understand and only by
reading further do I see what they mean by ’de-accumulated’. If working with
IFS PPFD is so hard, why didn’t the authors just calculate hourly (or minute-by-
minute) PPFD using cloud-cover and zenith angles?}
We have not used that ansatz, because we do not have the radiation fields in the relevant wavelengths available from OpenIFS output.
\end{itemize}
\newpage

### 2019-05-23 DryDep.tex
More work on the response. Almost there to have the first version ready and to start with the actual coading.

\subsubsection{Aerodynamic resistance}
\label{subsubsec:Ra}
The aerodynamic resistance is describing the turbulent transport of any substance down to the surface. %In \citet{WASP:Simpson2003,ACP:Simpson2012} it is described as
%\begin{equation}
%  R_a = \frac{1}{\kappa u_*}\left[{\ln{\left(\frac{z-d}{z_0}\right)}-\Psi_m\left(\frac{z-d}{L}\right)+\Psi_m\left(\frac{z_0}{L}\right)}\right],
%\end{equation}
%with the K\'{a}rm\'{a}n constant $\kappa = 0.40$, the friction velocity $u_*$, integrated stability equations for momentum $\Psi_m$,  a the deplacement height $d$ (typically $0.7\cdot h(N,\mathrm{lat})\,\unit{m}$ for vegetation other than forests), and the Obukhov length $L$. For certain values of $z$, $z_0$, and $L$, this may result in nonphysical (negative) values for $R_a$. For this reason, we diverge from the mOSaic scheme at this point and fall back to
For technical reasons, we used the method of \citet{Monteith1973}, wherein the sensible heat flux $\Phi_{Q_\text{sensible}}$ is written as
\begin{equation}
  \Phi_{Q_\text{sensible}} = \rho_\text{air}\,c_P \cdot \frac{\Delta T}{R_{a, H}}.
  \label{eq:sens_heat}
\end{equation}
Herein, $\rho_\text{air}$ is the air density, $c_P$ the specific heat at constant pressure $P$, $R_{a, H}$ is the diffusion resistance to sensible heat between surface and the reference height $z$, and $\Delta T$ is the difference between the surface temperature $T_0$ and the temperature at a reference height $T_z$. From the eddy diffusion theorem, a similar formulation can be derived
\begin{equation}
  \Phi_{Q_\text{sensible}} = \rho_\text{air}\,c_P \cdot \frac{\partial T}{\partial u} \cdot u_*^2.
  \label{eq:eddy_theo}
\end{equation}
For finite $z$, $\partial T \rightarrow \Delta T$. Comparing Eq.~(\ref{eq:sens_heat}) and Eq.~(\ref{eq:eddy_theo}) and assuming $R_{a, H} \equiv R_a$, we find the stability independent equation
\begin{equation}
  R_a = \frac{u_z}{u_*^2}.
\end{equation}
As reference height $z$, we use the height at mid-level of the lowermost model level (on average $10\,\unit{m}$). From the meteorological input fields $u_z$ is directly available, while $u_*$ can be computed via
\begin{equation}
  u_* = \sqrt{\frac{\tau_0}{\rho}},
\end{equation}
with the turbulent surface stress $\tau_0^2 = \tau_v^2 + \tau_u^2$, and northward turbulent surface stress $\tau_v$, eastward turbulent surface stress $\tau_u$.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Quasi-laminar layer resistance}
\label{subsubsec:Rb}
The quasi-laminar layer resistance $R_b^i$ is species specific and differs over land and ocean surfaces. In case a gridbox contains both, a weighted mean is calculated.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Over land, we use \citep[Eq.~(53),][]{ACP:Simpson2012}
\begin{equation}
  R_b^i = \frac{2}{\kappa u_*} \cdot \left(\frac{\text{Sc}_i}{\text{Pr}}\right)^{\frac{2}{3}},
  \label{eq:Rb}
\end{equation}
wherein $\text{Pr}$ is the Prandtl number (typically 0.72 for air and other gases) and $\text{Sc}_i$ is the Schmidt number for a gas $i$. Eq.~(\ref{eq:Rb}) differs from a similar formulation in \citet{ACP:Seinfeld2006} by a factor of roughly $1.25$. From $\text{Sc}_i = \nu/D_i$, with the kinematic viscosity of air $\nu$, we derive a Schmidt number in water equivalent:
\begin{equation}
  \text{Sc}_i = \frac{D_\chem{H_2O}}{D_i} \cdot \text{Sc}_\chem{H_2O},
\end{equation}
with the molecular diffusivity for any gas $D_i$, the Schmidt number of water ($\text{Sc}_\chem{H_2O}=0.6$) and its molecular diffusivity ($D_\chem{H_2O}=0.21\cdot10^{-4}\,\unit{m^2\,s^{-1}}$). The used ratios $D_\chem{H_2O}/D_i$ are taken from \citet[][Table S18]{ACP:Simpson2012}.\\

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsubsection*{Stomatal conductance}
The stomatal conductance is a measure of the rate of \chem{CO_2} exchange and evapotranspiration through the stomata of a leaf. There are several environmental conditions affecting the opening and closing of the stomata and hence their capability of respiration (e.g., light, available water, etc.). Stomata sluggishness, a state in which the stomata can no longer fully close, has been reported as ozone induced damage \citep{SR:Hoshika2015}, but is not taken into account in our formulation. To reflect part of the underlying mechanism, the stomatal conductance $G_\text{sto}$ in the EMAP scheme is computed using a multiplicative ansatz that is also explained in much detail in \citet{ICP:MappingManual2017}:
%
\begin{equation}
  G_\text{sto} = g^\text{m}_\text{max} \cdot f_\text{phen} \cdot f_\text{light} \cdot \max{\left\{f_\text{min}, f_T \cdot f_D \cdot f_\text{SW}\right\}}.
  \label{eq:stomatal}
\end{equation}
%
The factors herein are normalized and vary within the range $0-1$. They account for leaf phenology ($f_\text{phen}$), light ($f_\text{light}$), temperature ($f_T$), water vapor pressure deficit ($f_D$), and soil water content ($f_\text{SW}$).

@@

The non-stomatal conductance for \chem{O_3} consists of two terms, one depending on vegetation type and one depending on the soil/surface. For a land surface type $N$, we can write
\begin{equation}
  G_\text{ns}^\chem{O_3}(N) = \frac{\text{SAI}(N)}{r_\text{ext}} + \frac{1}{R_\text{inc}(N) + R_\text{gs}^\chem{O_3}(N)}.
  \label{eq:non-stomata-cond}
\end{equation}
$\text{SAI}(N)$ is the surface area index for vegetation type $N$, which is $\text{LAI}$ plus some value representing cuticles and others surfaces. The external leaf resistance is defined by
\begin{equation}
  r_\text{ext} = 2000\,\unit{s\,m^{-1}}\cdot F_T.
\end{equation}
Herein $F_T$ is a temperature correction factor for temperatures below $-1\,\unit{^\circ C}$ and $\{F_T \in \mathbb{R} | (1 \leq F_T \leq 2)\}$
\begin{equation}
  F_T = \exp{(-0.2\cdot(1+\theta_0))}.
\end{equation}
$\theta_0$ is the surface or $2\,\unit{m}$ temperature in \unit{^\circ C}. In general, $\text{SAI} \equiv \text{LAI}$ for all land types, except for: 
%
\begin{equation}
  \text{SAI} = 
  \begin{cases}
    \text{LAI}+1 & \emph{if} \quad \text{\emph{forest / wetland}},\\
    \text{LAI}\cdot 5/3.5 & \emph{if} \quad \text{\emph{cropland}, 1. part of growing season},\\
    \text{LAI} + 1.5 & \emph{if} \quad \text{\emph{cropland}, 2. part of growing season},\\
    0 & \emph{if} \quad \text{\emph{cropland}, winter}.
  \end{cases}
  \label{eq:SAI}
\end{equation}
%
Extending the mOSaic scheme to the southern hemisphere, the used growing season for crops is defined in Table~\ref{tab:growing_season}.
%
\begin{table}[t]
  \caption{Definition of growing season for crops used in the Oslo~CTM3 in northern hemisphere (NH) and southern hemisphere (SH).}
  \begin{tabular}{lrr}
    \tophline
    & 1. part & 2. part \\
    & (days) & (days) \\
    \middlehline
    NH & 90--140 & 141--270 \\
    SH & 272--322 & 323--452 \\
    \bottomhline
  \end{tabular}
  %\belowtable{} % Table Footnotes
  \label{tab:growing_season}
\end{table}
%

In this way, vegetation affects the conductance also by being there, not only by uptake through the stomata. The in-canopy resistance $R_\text{inc}$ is defined for each vegetated land type $N$ as
\begin{equation}
  R_\text{inc} = b \cdot \text{SAI}(N) \cdot \frac{h(N, \text{lat})}{u_*},
\end{equation}
where $h(N, \text{lat})$ is the latitude dependent vegetation height (see explanation at the end of this section) and $b = 14\,\unit{s^{-1}}$ is an empirical constant. According to \citet{ACP:Simpson2012}, the canopy resistance in EMEP does not take temperature and snow into account and is zero for non-vegetated surfaces.\\
$R_\text{gs}^\chem{O_3}(N)$ is tabulated and corrected for temperature by $F_T$ and snow cover fraction $f_\text{snow}$:
\begin{equation}
  \frac{1}{R_\text{gs}^\chem{O_3}(N)} = \frac{1-f_\text{snow}}{\hat{R}_\text{gs}^\chem{O_3}(N)} + \frac{f_\text{snow}}{R_\text{snow}^\chem{O_3}}.
\end{equation}
As initially mentioned, the necessary depth of snow to cover a certain type of vegetation differs. Therefore, we calculate a snow cover fraction $f_\text{snow}$ using the snow depth $S_D$, which is available in units of meter of water equivalent from the meteorological input data, scaled to $10\,\unit{\%}$ of the vegetation height.

@@

\subsubsection{Latitude dependent vegetation height}
The vegetation height $h(N, \text{lat})$ in the mOSaic scheme is linearly decreasing with latitude between $60\,\unit{^\circ}$ and $74\,\unit{^\circ N}$. To adapt this to a global model, we need to make a few additional assumptions. The tabulated height for each vegetation type $h(N)$ in the mOSaic scheme is regarded as constant at mid latitudes ($40\,\unit{^\circ}-60\,\unit{^\circ}$). Towards the poles, we decrease the height of each vegetation type using the same rate as described in \citet{ACP:Simpson2012}. At a latitude of $74\,\unit{^\circ}$ a minimum height of $3/10 \cdot h(N)$ is reached and kept constant. Towards the equator, we increase the height linearly so that at a latitude of $10\,\unit{^\circ}$ a maximum height of $2 \cdot h(N)$ is achieved which is then held constant. We also assume symmetry in both hemispheres. Presuming a typical tree height of $20\,\unit{m}$ at mid latitudes, this step-wise function yields a height of $8\,\unit{m}$ at high latitudes and $40\,\unit{m}$ in the tropics which is not unrealistic. For four example PFTs, results are shown in the Supplement (S.2, Fig.~S4).

@@

\begin{figure*}[t]
  \includegraphics[width=12cm]{fig02}
  \caption{Mapping of land surface categories. Either land surface categories from ISLSCP2 product of MODIS /  Community Land Model (CLM)~2 can be chosen for mapping to the nine land surface types used in the EMAP scheme. Water bodies and snow and ice categories of MODIS are actually not mapped. For both, MODIS and CLM~2 land surface categories, snow and ice cover is estimated from input meteorology, while water is defined as $1-\sum_{N=0}^{N_\text{max}} f_L(N)$. From MODIS category \emph{Barren or sparsely vegetated}, everything polward from $60\,\unit{^\circ}$ is defined as tundra, while everything equatorward is categorized as desert.}
  \label{fig:pft_mapping}
\end{figure*}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Pre-processing}
\label{subsec:pre-pro}
As mentioned in the previous section, there are two variables needed for computing the stomatal conductance which are not directly available from the meteorological input data. The greening season, as the time of the year in the mid and high latitudes when it is most likely for plants to grow, and the photosynthetic photon flux density, as the amount of light that plants need to photosynthesize. In the following, we present the necessary pre-processing of the variables. It is planed to implement an online computation of these variables into the Oslo~CTM3 later on.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Greening season}
\label{subsubsec:greening}
In Eqs.~(\ref{eq:non-stomata-cond}--\ref{eq:SAI}), \citet{ACP:Simpson2012} use prescribed start of growing season (SGS) and end of growing season (EGS) at $50\,\unit{^\circ N}$ ($d_\text{SGS}$, $d_\text{EGS}$) together with lapse rates ($\nabla d_\text{SGS}$, and $\nabla d_\text{EGS}$) to define phenology and dry deposition over agricultural areas. For the growing season of crops in the computation of non-stomatal conductances, we use also prescribed values (Table~\ref{tab:growing_season}), while for the stomatal conductances, as shown in Eq.~(\ref{eq:fphen}), we use the SGS and EGS derived parameters: day of greening season (GDAY), the time elapsed starting at the SGS, and the total length of the greening season (GLEN), the time span between EGS and SGS. Since the eurocentric EMEP parameterization of SGS and EGS is not applicable in a global model, another latitude dependent parameterization is needed. First, we have used a parameterization as implemented in the Sparse Matrix Operational Kernel Emissions -- Biogenic Emission Inventory System (SMOKE-BEIS; \href{https://www.epa.gov/air-emissions-modeling/biogenic-emission-inventory-system-beis}{model webpage}). SMOKE-BEIS has fixed values for SGS and EGS for all regions but NH mid latitudes ($23\,\unit{^\circ} < \text{lat} < 65\,\unit{^\circ}$), where it uses lapse rates of $\nabla d_\text{SGS} = 4.5$ and $\nabla d_\text{EGS} = 3.3$. As this parameterization is optimized for North America, it does not work well in Europe, e.g., most of northern Scandinavia has no allocated vegetation period.\\

>>>>

\subsubsection{Aerodynamic resistance}
\label{subsubsec:Ra}
{\bf ToDO: Implement \citet{AE:Erisman1994} formulation for Ra with d's from \citet{ACP:Simpson2012} and $\Psi_m$ from \citet{ACP:Seinfeld2006} and remove text about Monteith alternative!} The aerodynamic resistance is describing the turbulent transport of any substance down to the surface. Following \citet{AE:Garland1978}, \citet{WASP:Simpson2003,ACP:Simpson2012} has
\begin{equation}
  R_a = \frac{1}{\kappa u_*}\left[{\ln{\left(\frac{z-d}{z_0}\right)}-\Psi_m\left(\frac{z-d}{L}\right)+\Psi_m\left(\frac{z_0}{L}\right)}\right],
\end{equation}
with the K\'{a}rm\'{a}n constant $\kappa = 0.40$, the friction velocity $u_*$, integrated stability equations for momentum $\Psi_m$,  a the deplacement height $d$ (typically $0.7\cdot h(N,\mathrm{lat})\,\unit{m}$ for vegetation other than forests), and the Obukhov length $L$. For certain values of $z$, $z_0$, and $L$, this may result in nonphysical (negative) values for $R_a$. For this reason, we diverge from the mOSaic scheme at this point and fall back to
For technical reasons, we used the method of \citet{Monteith1973}, wherein the sensible heat flux $\Phi_{Q_\text{sensible}}$ is written as
\begin{equation}
  \Phi_{Q_\text{sensible}} = \rho_\text{air}\,c_P \cdot \frac{\Delta T}{R_{a, H}}.
  \label{eq:sens_heat}
\end{equation}
Herein, $\rho_\text{air}$ is the air density, $c_P$ the specific heat at constant pressure $P$, $R_{a, H}$ is the diffusion resistance to sensible heat between surface and the reference height $z$, and $\Delta T$ is the difference between the surface temperature $T_0$ and the temperature at a reference height $T_z$. From the eddy diffusion theorem, a similar formulation can be derived
\begin{equation}
  \Phi_{Q_\text{sensible}} = \rho_\text{air}\,c_P \cdot \frac{\partial T}{\partial u} \cdot u_*^2.
  \label{eq:eddy_theo}
\end{equation}
For finite $z$, $\partial T \rightarrow \Delta T$. Comparing Eq.~(\ref{eq:sens_heat}) and Eq.~(\ref{eq:eddy_theo}) and assuming $R_{a, H} \equiv R_a$, we find the stability independent equation
\begin{equation}
  R_a = \frac{u_z}{u_*^2}.
\end{equation}
As reference height $z$, we use the height at mid-level of the lowermost model level (on average $10\,\unit{m}$). From the meteorological input fields $u_z$ is directly available, while $u_*$ can be computed via
\begin{equation}
  u_* = \sqrt{\frac{\tau_0}{\rho}},
\end{equation}
with the turbulent surface stress $\tau_0^2 = \tau_v^2 + \tau_u^2$, and northward turbulent surface stress $\tau_v$, eastward turbulent surface stress $\tau_u$.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Quasi-laminar layer resistance}
\label{subsubsec:Rb}
The quasi-laminar layer resistance $R_b^i$ is species specific and differs over land and ocean surfaces. In case a gridbox contains both, a weighted mean is calculated.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Over land, we use \citep[Eq.~(53),][]{ACP:Simpson2012}
\begin{equation}
  R_b^i = \frac{2}{\kappa u_*} \cdot \left(\frac{\text{Sc}_i}{\text{Pr}}\right)^{\frac{2}{3}},
  \label{eq:Rb}
\end{equation}
wherein $\text{Pr}$ is the Prandtl number (typically 0.72 for air and other gases) and $\text{Sc}_i$ is the Schmidt number for a gas $i$. Eq.~(\ref{eq:Rb}) differs from a similar formulation in \citet{ACP:Seinfeld2006} by a factor of roughly $1.25$. From $\text{Sc}_i = \nu/D_i$, with the kinematic viscosity of air $\nu$, we derive a Schmidt number in water equivalent:
\begin{equation}
  \text{Sc}_i = \frac{D_\chem{H_2O}}{D_i} \cdot \text{Sc}_\chem{H_2O},
\end{equation}
with the molecular diffusivity for any gas $D_i$, the Schmidt number of water ($\text{Sc}_\chem{H_2O}=0.6$) and its molecular diffusivity ($D_\chem{H_2O}=0.21\cdot10^{-4}\,\unit{m^2\,s^{-1}}$). The used ratios $D_\chem{H_2O}/D_i$ are taken from \citet[][Table S18]{ACP:Simpson2012}.\\

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsubsection*{Stomatal conductance}
The stomatal conductance is a measure of the rate of \chem{CO_2} exchange and evapotranspiration through the stomata of a leaf. There are several environmental conditions affecting the opening and closing of the stomata and hence their capability of respiration (e.g., light, available water, etc.). Stomata sluggishness, a state in which the stomata can no longer fully close, has been reported as ozone induced damage \citep{SR:Hoshika2015}, but is not taken into account in our formulation. To reflect part of the underlying mechanism, the stomatal conductance $G_\text{sto}$ in the mOSaic scheme is computed using a multiplicative ansatz that is also explained in much detail in \citet{ICP:MappingManual2017}:
%
\begin{equation}
  G_\text{sto} = g^\text{m}_\text{max} \cdot f_\text{phen} \cdot f_\text{light} \cdot \max{\left\{f_\text{min}, f_T \cdot f_D \cdot f_\text{SW}\right\}}.
  \label{eq:stomatal}
\end{equation}
%
The factors herein are normalized and vary within the range $0-1$. They account for leaf phenology ($f_\text{phen}$), light ($f_\text{light}$), temperature ($f_T$), water vapor pressure deficit ($f_D$), and soil water content ($f_\text{SW}$).

@@

The non-stomatal conductance for \chem{O_3} consists of two terms, one depending on vegetation type and one depending on the soil/surface. For a land surface type $N$, we can write
\begin{equation}
  G_\text{ns}^\chem{O_3}(N) = \frac{\text{SAI}(N)}{r_\text{ext}} + \frac{1}{R_\text{inc}(N) + R_\text{gs}^\chem{O_3}(N)}.
  \label{eq:non-stomata-cond}
\end{equation}
$\text{SAI}(N)$ is the surface area index for vegetation type $N$, which is $\text{LAI}$ plus some value representing cuticles and others surfaces. The external leaf resistance is defined by
\begin{equation}
  r_\text{ext} = 2000\,\unit{s\,m^{-1}}\cdot F_T.
\end{equation}
Herein $F_T$ is a temperature correction factor for temperatures below $-1\,\unit{^\circ C}$ and $\{F_T \in \mathbb{R} | (1 \leq F_T \leq 2)\}$
\begin{equation}
  F_T = \exp{(-0.2\cdot(1+\theta_0))}.
\end{equation}
$\theta_0$ is the surface or $2\,\unit{m}$ temperature in \unit{^\circ C}. In general, $\text{SAI} \equiv \text{LAI}$ for all land types, except for: 
%
\begin{equation}
  \text{SAI} = 
  \begin{cases}
    \text{LAI}+1 & \emph{if} \quad \text{\emph{forest / wetland}},\\
    \text{LAI}\cdot 5/3.5 & \emph{if} \quad \text{\emph{cropland}, 1st part of growing season},\\
    \text{LAI} + 1.5 & \emph{if} \quad \text{\emph{cropland}, 2nd part of growing season},\\
    0 & \emph{if} \quad \text{\emph{cropland}, winter}.
  \end{cases}
  \label{eq:SAI}
\end{equation}
%
Extending the mOSaic scheme to the southern hemisphere, the used growing season for crops is defined in Table~\ref{tab:growing_season}.
%
\begin{table}[t]
  \caption{Definition of growing season for crops used in the Oslo~CTM3 in northern hemisphere (NH) and southern hemisphere (SH).}
  \begin{tabular}{lrr}
    \tophline
    & 1st part & 2nd part \\
    & (days) & (days) \\
    \middlehline
    NH & 90--140 & 141--270 \\
    SH & 272--322 & 323--452 \\
    \bottomhline
  \end{tabular}
  %\belowtable{} % Table Footnotes
  \label{tab:growing_season}
\end{table}
%

In this way, vegetation affects the conductance also by being there, not only by uptake through the stomata. The in-canopy resistance $R_\text{inc}$ \citep{AE:Erisman1994} is then modified with respect to each (vegetated) land type $N$
\begin{equation}
  R_\text{inc} = b \cdot \text{SAI}(N) \cdot \frac{h(N, \text{lat})}{u_*},
\end{equation}
where $h(N, \text{lat})$ is the latitude dependent vegetation height (see explanation at the end of this section) and $b = 14\,\unit{m^{-1}}$ is an empirical constant. According to \citet{ACP:Simpson2012}, the canopy resistance in EMEP does not take temperature and snow into account and is zero for non-vegetated surfaces.\\
$R_\text{gs}^\chem{O_3}(N)$ is tabulated and corrected for temperature by $F_T$ and snow cover fraction $f_\text{snow}$:
\begin{equation}
  \frac{1}{R_\text{gs}^\chem{O_3}(N)} = \frac{1-f_\text{snow}}{\hat{R}_\text{gs}^\chem{O_3}(N)} + \frac{f_\text{snow}}{R_\text{snow}^\chem{O_3}}.
\end{equation}
As initially mentioned, the necessary depth of snow to cover a certain type of vegetation differs. Therefore, we calculate a snow cover fraction $f_\text{snow}$ using the snow depth $S_D$, which is available in units of meter of water equivalent from the meteorological input data, scaled to $10\,\unit{\%}$ of the vegetation height.

@@

\subsubsection{Latitude dependent vegetation height}
The vegetation height $h(N, \text{lat})$ as decribed by \citet{ACP:Simpson2012} is linearly decreasing with latitude between $60\,\unit{^\circ}$ and $74\,\unit{^\circ N}$. To adapt this to a global model, we made a few additional assumptions. The tabulated height for each vegetation type $h(N)$ in the mOSaic scheme is regarded as constant at mid latitudes ($40\,\unit{^\circ}-60\,\unit{^\circ}$). Towards the poles, we decrease the height of each vegetation type using the same rate as described in \citet{ACP:Simpson2012}. At a latitude of $74\,\unit{^\circ}$ a minimum height of $3/10 \cdot h(N)$ is reached and kept constant. Towards the equator, we increase the height linearly so that at a latitude of $10\,\unit{^\circ}$ a maximum height of $2 \cdot h(N)$ is achieved which is then held constant. We also assume symmetry in both hemispheres. Presuming a typical tree height of $20\,\unit{m}$ at mid latitudes, this step-wise function yields a height of $8\,\unit{m}$ at high latitudes and $40\,\unit{m}$ in the tropics which is not unrealistic. For four example PFTs, results are shown in the Supplement (S.2, Fig.~S4).

@@

\begin{figure*}[t]
  \includegraphics[width=12cm]{fig02}
  \caption{Mapping of land surface categories. Either land surface categories from ISLSCP2 product of MODIS or the Community Land Model (CLM)~2 can be chosen for mapping to the nine land surface types usedwe use in the mOSaic scheme. Water bodies and snow and ice categories of MODIS are actually not mapped. For both, MODIS and CLM~2 land surface categories, snow and ice cover is estimated from input meteorology, while water is defined as $1-\sum_{N=0}^{N_\text{max}} f_L(N)$. From MODIS category \emph{Barren or sparsely vegetated}, everything polward from $60\,\unit{^\circ}$ is defined as tundra, while everything equatorward is categorized as desert.}
  \label{fig:pft_mapping}
\end{figure*}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Pre-processing}
\label{subsec:pre-pro}
As mentioned in the previous section, there are two variables needed for computing the stomatal conductance which are not directly available from the meteorological input data. The greening season, as the time of the year in the mid and high latitudes when it is most likely for plants to grow, and the photosynthetic photon flux density, as the amount of light that plants need to photosynthesize. In the following, we present the necessary pre-processing of the variables. It is planed to implement an online computation of these variables into the Oslo~CTM3 later on.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Greening season}
\label{subsubsec:greening}
In Eqs.~(\ref{eq:non-stomata-cond}--\ref{eq:SAI}), \citet{ACP:Simpson2012} use prescribed start of growing season (SGS) and end of growing season (EGS) at $50\,\unit{^\circ N}$ ($d_\text{SGS}$, $d_\text{EGS}$) together with lapse rates ($\nabla d_\text{SGS}$, and $\nabla d_\text{EGS}$) to define phenology and dry deposition over agricultural areas. For the growing season of crops in the computation of non-stomatal conductances, we use also prescribed values (Table~\ref{tab:growing_season}), while for the stomatal conductances, as shown in Eq.~(\ref{eq:fphen}), we use the SGS and EGS derived parameters: day of greening season (GDAY), the time elapsed starting at the SGS, and the total length of the greening season (GLEN), the time span between EGS and SGS. Since the eurocentric parameterization of SGS and EGS in \citet{ACP:Simpson2012} is not applicable in a global model, another latitude dependent parameterization is needed. First, we have used a parameterization as implemented in the Sparse Matrix Operational Kernel Emissions -- Biogenic Emission Inventory System (SMOKE-BEIS; \href{https://www.epa.gov/air-emissions-modeling/biogenic-emission-inventory-system-beis}{model webpage}). SMOKE-BEIS has fixed values for SGS and EGS for all regions but NH mid latitudes ($23\,\unit{^\circ} < \text{lat} < 65\,\unit{^\circ}$), where it uses lapse rates of $\nabla d_\text{SGS} = 4.5$ and $\nabla d_\text{EGS} = 3.3$. As this parameterization is optimized for North America, it does not work well in Europe, e.g., most of northern Scandinavia has no allocated vegetation period.\\
